---
title: '1. Data Preparation'
author: "Sugiyama, Takeshi"
output:
  html_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)

```
-----

The following code chunk will:

(1) read-in the abalones dataset, defining a new data frame, "mydata,"
(2) return the structure of that data frame, and
(3) calculate new variables, VOLUME and RATIO.

```{r analysis_setup1, message = FALSE, warning = FALSE}

# 1) Use read.csv() to read the abalones.csv into R, assigning the data frame to "mydata."

mydata <- read.csv("abalones.csv", sep = ",", stringsAsFactors = TRUE)


# 2) Use the str() function to verify the structure of "mydata." We should have 1036 observations
# of eight variables.

str(mydata)

# 3) Define two new variables, VOLUME and RATIO. Use the following statements to define VOLUME and
# RATIO as variables appended to the data frame "mydata."

mydata$VOLUME <- mydata$LENGTH * mydata$DIAM * mydata$HEIGHT
mydata$RATIO <- mydata$SHUCK / mydata$VOLUME

```

-----

***##### Section 1: Summarizing the data.***

```{r Part_1a}
summary(mydata)

table(mydata$CLASS, mydata$RINGS)

```

The variables being discussed here is SEX and CLASS. SEX contains male, female, and infant. CLASS are ordered age classification based on RINGS (A1=youngest, and A6= oldest). Overall, the distribution seems like it's following a normal distribution, but with a slight left skewness.
Potential outliers may be abalones that are considered infants in the A4 and A5 groups. Since A4 and A5 are older, their sex should have been identified and it could be an error in the data.

```{r Part_1b}
t1 <- table(mydata$SEX, mydata$CLASS)
addmargins(t1)

barplot(t1, legend = TRUE, main="Overall Distribution")
par(mfrow=c(1,3))
barplot(t1["F",], main="F")
barplot(t1["I",], main="I")
barplot(t1["M",], main="M")

```

**Essay Question (2 points):  Discuss the sex distribution of abalones.  What stands out about the distribution of abalones by CLASS?**

***Answer: (Enter your answer here.)***
Female and male abalones are left skewed and infant abalones are right skewed in the ordered class. This is most likely because it is harder to identify the sex of younger abalones. The number of infant abalones decrease as the age groups increase because the sex are identified in older groups. 

(1)(c) (1 point) Select a simple random sample of 200 observations from "mydata" and identify this sample as "work." Use *set.seed(123)* prior to drawing this sample. Do not change the number 123. Note that *sample()* "takes a sample of the specified size from the elements of x." We cannot sample directly from "mydata." Instead, we need to sample from the integers, 1 to 1036, representing the rows of "mydata." Then, select those rows from the data frame (Kabacoff Section 4.10.5 page 87). 

Using "work", construct a scatterplot matrix of variables 2-6 with *plot(work[, 2:6])* (these are the continuous variables excluding VOLUME and RATIO). The sample "work" will not be used in the remainder of the assignment. 

```{r Part_1c}
set.seed(123)
sample_ints <-sample(1:1036, 200)
work <- mydata[sample_ints, ]
plot(work[, 2:6])

```

-----

***##### Section 2: (5 points) Summarizing the data using graphics.***

(2)(a) (1 point) Use "mydata" to plot WHOLE versus VOLUME. Color code data points by CLASS.

```{r Part_2a}
class_color <- c("lightgreen", "darkgreen", "orange", "red", "blue")
plot(mydata$WHOLE, mydata$VOLUME, col=class_color[mydata$CLASS], xlab = "Whole", ylab="Volume")
legend("topleft", legend=sort(unique(mydata$CLASS)),col=class_color, lwd=1, bty="o")
```


(2)(b) (2 points) Use "mydata" to plot SHUCK versus WHOLE with WHOLE on the horizontal axis. Color code data points by CLASS. As an aid to interpretation, determine the maximum value of the ratio of SHUCK to WHOLE.  Add to the chart a straight line with zero intercept using this maximum value as the slope of the line. If you are using the 'base R' *plot()* function, you may use *abline()* to add this line to the plot. Use *help(abline)* in R to determine the coding for the slope and intercept arguments in the functions. If you are using ggplot2 for visualizations, *geom_abline()* should be used.

```{r Part_2b}
shuck_whole_ratio <- mydata$SHUCK / mydata$WHOLE
slope <- shuck_whole_ratio[which.max(shuck_whole_ratio)]
class_color <- c("lightgreen", "darkgreen", "orange", "red", "blue")
plot(mydata$WHOLE, mydata$SHUCK,  col=class_color[mydata$CLASS], xlab = "Whole", ylab="Shuck")
legend("topleft", legend=sort(unique(mydata$CLASS)),col=class_color, lwd=1, bty="o")
abline(a=0, b=slope)

```

**Essay Question (2 points):  How does the variability in this plot differ from the plot in (a)?  Compare the two displays.  Keep in mind that SHUCK is a part of WHOLE.  Consider the location of the different age classes.**

***Answer: (Enter your answer here.)***
Both plots have a positive correlation to whole. As age groups increase, whole increases and the y variable of volume and shuck increase as well. 
The first plot has A3~A5 scattered around the middle and top right part of the graph. This behavior seems random without a clear trend.
However, for the second plot, we can observe certain trends. The younger groups such as A1~A3 are much closer to the maximum line compared to the older groups such A4~A5. It seems like as abalones age, the ratio of shuck to whole starts to decrease. This is an interesting finding because the first plot is a visualization of volume vs whole. Even though the volume may increase as abalones age, that doesn't mean the shuck will increase as well. 
-----

***### Section 3: (8 points) Getting insights about the data using graphs.***

(3)(a) (2 points) Use "mydata" to create a multi-figured plot with histograms, boxplots and Q-Q plots of RATIO differentiated by sex. This can be done using *par(mfrow = c(3,3))* and base R or *grid.arrange()* and ggplot2. The first row would show the histograms, the second row the boxplots and the third row the Q-Q plots. Be sure these displays are legible.  

```{r Part_3a}
par(mfcol = c(3,3))
for (sex in unique(mydata$SEX)){
  temp_data <- mydata[mydata$SEX == sex, "RATIO"]
  hist(temp_data, main = sprintf("Histogram: Sex=%s", sex))
  boxplot(temp_data, main = sprintf("Boxplot: Sex=%s", sex))
  qqnorm(temp_data, pch = 1, frame = FALSE,  main = sprintf("QQPlot: Sex=%s", sex))
  qqline(temp_data, lwd = 2)
rm(temp_data)
}
```

**Essay Question (2 points): Compare the displays.  How do the distributions compare to normality? Take into account the criteria discussed in the sync sessions to evaluate non-normality.**

***Answer: (Enter your answer here.)***
All three sexes are close to a normality. However, all 3 sexes seem to be right skewed as we can observe from the histogram and qqplot. The boxplot shows there are outliers in all 3 sexes above the upper extreme. 


(3)(b) (2 points) The boxplots in (3)(a) indicate that there are outlying RATIOs for each sex. *boxplot.stats()* can be used to identify outlying values of a vector. Present the abalones with these outlying RATIO values along with their associated variables in "mydata". Display the observations by passing a data frame to the kable() function. Basically, we want to output those rows of "mydata" with an outlying RATIO, but we want to determine outliers looking separately at infants, females and males.

```{r Part_3b}
outlier_index_vec <- c()
for (sex in unique(mydata$SEX)){
  temp_outliers <- boxplot.stats(mydata[mydata$SEX == sex, "RATIO"])$out
  outlier_idx <- which((mydata$RATIO %in% temp_outliers) & (mydata$SEX == sex))
  outlier_index_vec <- append(outlier_index_vec, outlier_idx)
rm(temp_outliers, outlier_idx)
}
kable(mydata[outlier_index_vec,])
```

**Essay Question (2 points):  What are your observations regarding the results in (3)(b)?**

***Answer: (Enter your answer here.)***
The majority of the outliers have a ratio above 0.22. There was 1 observation that had a ratio of 0.06 which made it as an outlier in the lower extreme. Also, there were no outliers for abalones in the A4 class. 
-----

***### Section 4: (8 points) Getting insights about possible predictors.***

(4)(a) (3 points) With "mydata," display side-by-side boxplots for VOLUME and WHOLE, each differentiated by CLASS There should be five boxes for VOLUME and five for WHOLE. Also, display side-by-side scatterplots:  VOLUME and WHOLE versus RINGS. Present these four figures in one graphic:  the boxplots in one row and the scatterplots in a second row. Base R or ggplot2 may be used.


```{r Part_4a}
par(mfrow=c(2,2))
boxplot(mydata$VOLUME~mydata$CLASS, xlab="Class", ylab="Volume")
boxplot(mydata$WHOLE~mydata$CLASS, xlab="Class", ylab="Whole")
plot(mydata$VOLUME, mydata$RINGS, xlab="Volume", ylab="Rings")
plot(mydata$WHOLE, mydata$RINGS, xlab="Whole", ylab="Rings")
```

**Essay Question (5 points) How well do you think these variables would perform as predictors of age?  Explain.**

***Answer: (Enter your answer here.)***
Since there is a positive correlation between these variables, I think it would do fairly well to predict the age. However, as age approaches the A4 & A5 class, it would be difficult to classify because both groups have similar boxplot stats for volume and whole.
-----

***### Section 5: (12 points) Getting insights regarding different groups in the data.***

(5)(a) (2 points) Use *aggregate()* with "mydata" to compute the mean values of VOLUME, SHUCK and RATIO for each combination of SEX and CLASS. Then, using *matrix()*, create matrices of the mean values. Using the "dimnames" argument within *matrix()* or the *rownames()* and *colnames()* functions on the matrices, label the rows by SEX and columns by CLASS. Present the three matrices (Kabacoff Section 5.6.2, p. 110-111). The *kable()* function is useful for this purpose.  You do not need to be concerned with the number of digits presented.

```{r Part_5a}
cols <-c("VOLUME", "SHUCK", "RATIO")
aggs_means <- aggregate(mydata[,cols], by=list(mydata$SEX, mydata$CLASS), FUN=mean)
dim_names <- list(c("F", "I", "M"), c("A1", "A2", "A3", "A4", "A5"))
volume_mat <- matrix(aggs_means[, 3], nrow=3, ncol = 5, dimnames = dim_names)
shuck_mat <- matrix(aggs_means[, 4], nrow=3, ncol = 5, dimnames = dim_names)
ratio_mat <- matrix(aggs_means[, 5], nrow=3, ncol = 5, dimnames = dim_names)
kable(volume_mat)
kable(shuck_mat)
kable(ratio_mat)
```

(5)(b) (3 points) Present three graphs. Each graph should include three lines, one for each sex. The first should show mean RATIO versus CLASS; the second, mean VOLUME versus CLASS; the third, mean SHUCK versus CLASS. This may be done with the 'base R' *interaction.plot()* function or with ggplot2 using *grid.arrange()*.

```{r Part_5b, fig.width = 9}
gender_color <- c("red", "purple", "blue")
par(mfrow=c(3,1))
for (v in c("RATIO", "VOLUME", "SHUCK")){
  interaction.plot(x.factor = mydata$CLASS, #x-axis variable
                   trace.factor = mydata$SEX, #variable for lines
                   response = mydata[,v], #y-axis variable
                   fun = mean, #metric to plot
                   ylab = sprintf("%s", v),
                   xlab = "CLASS",
                   col = gender_color,
                   lty = 1, #line type
                   lwd = 3, #line width
                   trace.label = "Gender")
}

```

**Essay Question (2 points):  What questions do these plots raise?  Consider aging and sex differences.**

***Answer: (Enter your answer here.)***
It seems volume and shuck follow a similar trend. As age increases, the volume and shuck increases as well. The females have the highest average volume and shuck in any age group for these 2 variables. 
However, when it comes to ratio, it follows a different trend. As age increases, the average ratio decreases. Also, the lines intersect frequently. In A1 groups, the infants have the highest ratio. After A2, the female and male keeps intersecting. 
I wonder why the trend for the ratio are inconsistent compared to volume and shuck. It makes me even wonder if some of the abalones were correctly classified because there are inconsistencies in the ratio. 

5(c) (3 points) Present four boxplots using *par(mfrow = c(2, 2)* or *grid.arrange()*. The first line should show VOLUME by RINGS for the infants and, separately, for the adult; factor levels "M" and "F," combined. The second line should show WHOLE by RINGS for the infants and, separately, for the adults. Since the data are sparse beyond 15 rings, limit the displays to less than 16 rings. One way to accomplish this is to generate a new data set using subset() to select RINGS < 16.  Use ylim = c(0, 1100) for VOLUME and ylim = c(0, 400) for WHOLE.  If you wish to reorder the displays for presentation purposes or use ggplot2 go ahead.

```{r Part_5c}
library(dplyr)
newdata <- subset(mydata, RINGS < 16)
newdata <- newdata %>% 
  mutate(newdata, "ADULT":= (Adult = if_else(SEX == "I", "Infant", "Adult")))
infant_mask <- which(newdata$ADULT == "Infant")
adult_mask <- which(newdata$ADULT == "Adult")
par(mfrow = c(2, 2))
boxplot(newdata[infant_mask,]$VOLUME~newdata[infant_mask,]$RINGS, xlab="Rings (Infant)", ylab="Volume", ylim = c(0, 1100))
boxplot(newdata[adult_mask,]$VOLUME~newdata[adult_mask,]$RINGS, xlab="Rings (Adult)", ylab="Volume", ylim = c(0, 1100))
boxplot(newdata[infant_mask,]$WHOLE~newdata[infant_mask,]$RINGS, xlab="Rings (Infant)", ylab="WHOLE", ylim = c(0, 400))
boxplot(newdata[adult_mask,]$WHOLE~newdata[adult_mask,]$RINGS, xlab="Rings (Adult)", ylab="WHOLE", ylim = c(0, 400))
```

**Essay Question (2 points):  What do these displays suggest about abalone growth?  Also, compare the infant and adult displays.  What differences stand out?**

***Answer: (Enter your answer here.)***
We can suggest that as abalones age, the volume and whole increase. Comparing the infant and adults, we can observe that the adults have more variability as rings increases. The infants have a relatively small variance compared to the adults. 

-----

***### Section 6: (11 points) Conclusions from the Exploratory Data Analysis (EDA).***


**Conclusions**


**Essay Question 1) (5 points)	Based solely on these data, what are plausible statistical reasons that explain the failure of the original study? Consider to what extent physical measurements may be used for age prediction.**

***Answer: (Enter your answer here.)***
The biggest reason that the original study failed was because it was not possible to completely predict the age of the abalone from the physical measurements. Specifically, the age group of A4~A5 had identical sizes which was difficult to differentiate. For the age groups in early ages such as A1~A3, it may be possibly to predict the age with the physical appearances. 

**Essay Question 2) (3 points) Do not refer to the abalone data or study.  If you were presented with an overall histogram and summary statistics from a sample of some population or phenomenon and no other information, what questions might you ask before accepting them as representative of the sampled population or phenomenon?**

***Answer: (Enter your answer here.)***
I would inquire if there are certain groups that need to be considered. For example, for studies related to health, age is an extremely important variable. During the COVID pandemic, the Simpson's paradox became a topic of concern while analyzing the statistics of fatality rate and vaccine effectiveness.  https://www.lancaster.ac.uk/stor-i-student-sites/robyn-goldsmith/2021/04/02/being-careful-with-covid-19-statistics-simpsons-paradox/ 
I would inquire about any demographic information as possible. 

**Essay Question 3)  (3 points)	Do not refer to the abalone data or study.  What do you see as difficulties analyzing data derived from observational studies? Can causality be determined?  What might be learned from such studies?**

***Answer: (Enter your answer here.)***
One of the most difficult elements of observational studies is that the sample may not represent the population. For example, even though a study used stratified sampling methods to gather data, the same results may not hold true if the data was collected in an entirely different environment (e.g., country, time of day/year, etc.). With observational studies, correlation can possibly be determined, but not causality. Determining causality would be extremely difficult and it will require domain knowledge and critical thinking. 