---
title: "EDA"
author: "Sugiyama, Takeshi"
output:
  html_document: default
---

```{r 2setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)

```

## 2. Data Analysis

-----

The structure of the data. We can see that there are 1,036 observations (rows) and 10 variables (columns).

```{r 2analysis_setup, message = FALSE, warning = FALSE}

library(dplyr)
library(flux)
library(ggplot2)
library(gridExtra)
library(knitr)
library(rockchalk)
library(tidyverse)
library(glue) 

mydata <- read.csv("mydata.csv", sep = ",", stringsAsFactors = TRUE)
str(mydata)

```
```{r 2Part_1a, fig.width = 12}
my_layout <- matrix(c(1,2,3,3), ncol=2, byrow=TRUE)
layout(my_layout)
hist(x = mydata$RATIO, main="Abalone RATIO Frequencies", col="orange")
qqnorm(mydata$RATIO, col="orange")
qqline(mydata$RATIO)
boxplot(mydata$RATIO~mydata$CLASS, col="orange")

s <- rockchalk::skewness(x=mydata$RATIO)
k <- rockchalk::kurtosis(x=mydata$RATIO, excess=FALSE)
k_ex <- rockchalk::kurtosis(x=mydata$RATIO, excess=TRUE)
glue("skewness: {s}")
glue("kurtosis: {k}")
glue("excess kurtosis: {k_ex}")
```


```{r 2Part_1b, fig.width = 12, fig.height = 8}

mydata$L_RATIO <- log10(mydata$RATIO)

my_layout <- matrix(c(1,2,3,3), ncol=2, byrow=TRUE)
layout(my_layout)
hist(x = mydata$L_RATIO, main="Abalone L_RATIO Frequencies", col="gold")
qqnorm(mydata$L_RATIO, col="gold")
qqline(mydata$L_RATIO)
boxplot(mydata$L_RATIO~mydata$CLASS, col="gold")


s <- rockchalk::skewness(x=mydata$L_RATIO)
k <- rockchalk::kurtosis(x=mydata$L_RATIO, excess=FALSE)
k_ex <- rockchalk::kurtosis(x=mydata$L_RATIO, excess=TRUE)
glue("skewness: {s}")
glue("kurtosis: {k}")
glue("excess kurtosis: {k_ex}")

```

```{r Part_1c}
bartlett.test(mydata$RATIO~mydata$CLASS)
bartlett.test(mydata$L_RATIO~mydata$CLASS)

```

L_RATIO (log ratio) exhibits a better conformance to a normal distribution with homogeneous variances across age. The skewness is close to 0 which indicates it's a symmetrical distribution. The kurtosis is also close to 3 which indicates it is mesokurtic and resembles a normal distribution peak. We can visually see this with the histogram and qqplot as well. We can also see that the L_RATIO have homogeneous variances across ages with the Barlett test. Since the p-value is much greater than 0.05, we fail to reject the null hypothesis of having identical variances. On the other hand, the RATIO has p-value of 0.0002531 for the Barlett test which indicates that the variances are not homogeneous across ages.


```{r 2Part_2a}
mod1 <- aov(L_RATIO~CLASS*SEX, data = mydata)
summary(mod1)

mod2 <- aov(L_RATIO~CLASS+SEX, data = mydata)
summary(mod2)
```
While we get statistically significant results with CLASS and SEX independently, that is not the case when we factor CLASS and SEX. We can say there is an interaction effect between these two independent variables. Since both of these variables affects the outcome of L_RATIO, we inevitably get a non-significant result when these two are interacting.

```{r 2Part_2b}
TukeyHSD(mod2)

```

We can identify many statistically significant results from different age groups. The only non-significant result we observe was from A2-A1. This indicates that as the abolone's age group change (besides from A1 to A2), the L_RATIO will change as well. Male and female abalones can be combined into a single category because both have significant differences from infants and we have an extremely high adjusted p-value which suggests there is no difference between male and female.

```{r 2Part_3a1}
# Here, we show how to define the new variable TYPE using only base R functions:

mydata$TYPE <- factor(ifelse(mydata$SEX == "I", "I", "ADULT"))
table(mydata$TYPE)

```

```{r 2Part_3a2, fig.width = 12}
ggplot(data = mydata, aes(x = VOLUME)) +
  geom_histogram(fill="red") +
  facet_grid(. ~TYPE)
```

Infants have a right skewed distribution compared to adults. It would be hard to separate infants and adults around the lower volume areas, since there are a lot of overlaps in the two distributions (especially volumes under 300). For larger volumes, it may be easier to separate infants and adults. For volumes over 500, most of the abalones tend to be adults.

```{r 2Part_3b, fig.width = 12, fig.height = 8}

mydata$L_SHUCK <- log10(mydata$SHUCK)
mydata$L_VOLUME <- log10(mydata$VOLUME)

grid.arrange(
  ggplot(data = mydata, aes(x = VOLUME, y = SHUCK, col = CLASS)) +
    geom_point(alpha = 0.5) + theme(legend.position = "bottom"),
  ggplot(data = mydata, aes(x = L_VOLUME, y = L_SHUCK, col = CLASS)) +
    geom_point(alpha = 0.5) + theme(legend.position = "bottom"), 
  ggplot(data = mydata, aes(x = VOLUME, y = SHUCK, col = TYPE)) +
    geom_point(alpha = 0.5) + theme(legend.position = "bottom"), 
  ggplot(data = mydata, aes(x = L_VOLUME, y = L_SHUCK, col = TYPE)) +
    geom_point(alpha = 0.5) + theme(legend.position = "bottom"),
  nrow=2
)

```

The log-transformation reduces the variability of the relationship between volume and shuck we can make a better distinction between groups. We can see that there is a positive correlation between volume and shuck. For class levels, we can observe that the younger age groups, especially the A1 groups, are separated from the rest. We can observe that in general, when age groups increase, the shuck and volume increase as well. This observation is replicated in the type level. Most of the infants are scattered around the lower end of the volume and shuck. The adults are scattered around the higher end of the volume and shuck.

```{r 2Part_4a1}
mydata$TYPE[mydata$CLASS == "A4" | mydata$CLASS == "A5"] <- "ADULT"
table(mydata$TYPE)

```


```{r 2Part_4a2}
m_reg1 <- lm(L_SHUCK~L_VOLUME + CLASS + TYPE, data = mydata)
summary(m_reg1)
```

The summary here shows the reference level of class as A1 and type as adult. The coefficients for classes A2-A5 are the differences associated with A1 and adult. We can imply that the classes A1-A5 are highly correlated variables. When we plotted the data in the scatter plot earlier, it showed positive correlation for volume and shuck for all classes A1-A5 independently. However, when we fit the model and considered all the predictor variables together, we can see that the classes A2-A5 have a negative effect. The coefficients shows that if the data belongs to a  class in A2-A5 and holding every other predictors constant, it will have a negative effect when the reference class is A1 and reference type is adult.

Compared to class, type is not as an important predictor for shuck. The coefficient are much smaller compared to most of the other classes. However, it is important to keep in mind that the dependent variable is in log 10 scale. This means that even though it may be a small coefficient, it still can influence the results greatly.

-----

```{r 2Part_5am, fig.width = 12}
my_layout <- matrix(c(1,2), ncol=2, byrow=TRUE)
layout(my_layout)
hist(x = m_reg1$residuals, main="Residuals of Model", col="lightblue")
qqnorm(m_reg1$residuals, col="lightblue")
qqline(m_reg1$residuals)

s <- rockchalk::skewness(x=m_reg1$residuals)
k <- rockchalk::kurtosis(x=m_reg1$residuals, excess=FALSE)
k_ex <- rockchalk::kurtosis(x=m_reg1$residuals, excess=TRUE)
glue("skewness: {s}")
glue("kurtosis: {k}")
glue("excess kurtosis: {k_ex}")
```


```{r 2Part_5b, fig.width = 12, fig.height = 8}
tempdf <- data.frame(L_VOLUME = mydata$L_VOLUME, CLASS = mydata$CLASS, TYPE = mydata$TYPE, residuals = m_reg1$residuals)
grid.arrange(
  ggplot(data = tempdf, aes(x = L_VOLUME, y = residuals, col = CLASS)) +
    geom_point(alpha = 0.5) + theme(legend.position = "bottom"),
  ggplot(data = tempdf, aes(x = L_VOLUME, y = residuals, col = TYPE))  +
    geom_point(alpha = 0.5) + theme(legend.position = "bottom"), 
  ggplot(data = tempdf, aes(x = CLASS, y = residuals))  +
    geom_boxplot() + theme(legend.position = "bottom"),
  ggplot(data = tempdf, aes(x = TYPE, y = residuals))  +
    geom_boxplot() + theme(legend.position = "bottom"), 
  nrow=2
)
bartlett.test(tempdf$residuals~tempdf$CLASS)
bartlett.test(tempdf$residuals~tempdf$TYPE)
```

With the above calculations, we checked the assumptions of linear regression. We checked if the residuals were normally distributed and  tested the homogeneity of variance of the residuals across classes. Both of these assumptions were met. We can also observe that volume and shuck have a linear relationship from 3(b) and we know each abalones are independent from each other. Therefore, this model indeed fit the dataset and it may be useful for harvesting decisions.

-----

Harvest Strategy:

There is a tradeoff faced in managing abalone harvest. The infant population must be protected since it represents future harvests. On the other hand, the harvest should be designed to be efficient with a yield to justify the effort. This assignment will use VOLUME to form binary decision rules to guide harvesting. If VOLUME is below a "cutoff" (i.e. a specified volume), that individual will not be harvested. If above, it will be harvested. Different rules are possible.The Management needs to make a decision to implement 1 rule that meets the business goal.

-----

```{r 2Part_6a}

idxi <- mydata$TYPE == "I"
idxa <- mydata$TYPE == "ADULT"

max.v <- max(mydata$VOLUME)
min.v <- min(mydata$VOLUME)
delta <- (max.v - min.v)/10000
prop.infants <- numeric(10000)
prop.adults <- numeric(10000)
volume.value <- numeric(10000)

total.infants <- sum(idxi)  
total.adults <- sum(idxa)

for (k in 1:10000) { 
	value <- min.v + k*delta
	volume.value[k] <- value
	prop.infants[k] <- sum(mydata$VOLUME[idxi] <= value)/total.infants
	prop.adults[k] <-  sum(mydata$VOLUME[idxa] <= value)/total.adults
}

```


```{r 2Part_6b}
# Largest infant volume
(max_inf_vol <- max(mydata$VOLUME[mydata$TYPE == "I"]))  # [1] 526.6383

# Proportion of adults harvested
sum(mydata$VOLUME[mydata$TYPE == "ADULT"] > max_inf_vol) /
  total.adults  # [1] 0.2476573

# Add code to calculate the proportion of infants harvested
sum(mydata$VOLUME[mydata$TYPE == "I"] > max_inf_vol) /
  total.infants  # [1] 0.0

# If we use the largest infant volume, we harvest approximately 24.8% of adults and 0%,
# as expected, of infants.

```


```{r 2Part_6c}
# Add code to determine the median infant volume:
(med_inf_vol <- median(mydata$VOLUME[mydata$TYPE == "I"])) # [1] 133.8214

# Add code to calculate the proportion of infants harvested
sum(mydata$VOLUME[mydata$TYPE == "I"] > med_inf_vol) /
  total.infants # [1] 0.4982699

# Add code to calculate the proportion of adults harvested
sum(mydata$VOLUME[mydata$TYPE == "ADULT"] > med_inf_vol) /
  total.adults # [1] 0.9330656

# If we use the median infant volume as our cutoff, we harvest almost 50% of our infants
# and a little more than 93% of our adults.


# Add code to determine the median adult volume:
(med_adu_vol <- median(mydata$VOLUME[mydata$TYPE == "ADULT"])) # [1] 384.5584

# Add code to calculate the proportion of infants harvested
sum(mydata$VOLUME[mydata$TYPE == "I"] > med_adu_vol) /
  total.infants # [1] 0.02422145

# Add code to calculate the proportion of adults harvested
sum(mydata$VOLUME[mydata$TYPE == "ADULT"] > med_adu_vol) /
  total.adults # [1] 0.4993307

# If we use the median adult volume as our cutoff, we harvest almost 50% of adults
# and approximately 2.4% of infants.

```

```{r 2Part_6d, fig.width = 12, fig.height = 6}

library(reshape2) # install.packages("reshape2")

harvest <- data.frame(Adults = prop.adults, Infants = prop.infants, Volume = volume.value)
harvest <- melt(harvest, id.vars="Volume", variable.name = "Type", value.name="Proportion")

ggplot(data = harvest, aes(x = Volume, y = Proportion, col = Type)) +
  geom_line() +
  geom_vline(xintercept = max_inf_vol, col='green') + 
  geom_text(aes(x=max_inf_vol+20, label=glue("Protect All Infants: {round(max_inf_vol, 1)}"), y=1.05), colour="green") +
  geom_vline(xintercept = med_inf_vol, col='red') + 
  geom_text(aes(x=med_inf_vol-20, label=glue("Median Infants: {round(med_inf_vol, 1)}"), y=1.05), colour="red") +
  geom_vline(xintercept = med_adu_vol, col='orange') + 
  geom_text(aes(x=med_adu_vol-20, label=glue("Median Adults: {round(med_adu_vol, 1)}"), y=1.05), colour="orange") + 
  theme(legend.position = "bottom")

```

We can clearly observe a difference in median between infants and adults. There will be a trade off of trying to preserve as much infants versus trying to harvest as much adults as possible. Since we want to save as much infants as possible, we don't want to choose the median infant point as the cut off. This would harvest roughly half of all infants which would have a negative effect on future harvests. Choosing the median adult point as the cut off would be the better option because it would still save roughly 97.6% of all infants while harvesting half of all adults. There needs to be a business decision made based on the choosing the right cut off point for the business.

-----


More harvest strategies:

This part will address the determination of a cutoff volume.value corresponding to the observed maximum difference in harvest percentages of adults and infants. In other words, we want to find the volume value such that the vertical distance between the infant curve and the adult curve is maximum. To calculate this result, the vectors of proportions from item (6) must be used. These proportions must be converted from "not harvested" to "harvested" proportions by using (1 - prop.infants) for infants, and (1 - prop.adults) for adults. The reason the proportion for infants drops sooner than adults is that infants are maturing and becoming adults with larger volumes.

-----

```{r 2Part_7a}
difference <- ((1 - prop.adults) - (1 - prop.infants))
diff_peak <- volume.value[which.max(difference)]
plot(x = volume.value, y = difference, type="l")
abline(v=max_inf_vol, col="green")
abline(v=med_inf_vol, col="red")
abline(v=med_adu_vol, col="orange")
abline(v=diff_peak, col="black")
text(x=diff_peak, y=0, glue("Difference Peak: {round(diff_peak, 1)}"))
```

```{r 2Part_7b}

y.loess.a <- loess(1 - prop.adults ~ volume.value, span = 0.25,
	family = c("symmetric"))
y.loess.i <- loess(1 - prop.infants ~ volume.value, span = 0.25,
	family = c("symmetric"))
smooth.difference <- predict(y.loess.a) - predict(y.loess.i)
smoothdiff_peak <- volume.value[which.max(smooth.difference)]
plot(x = volume.value, y = smooth.difference, type="l")
abline(v=max_inf_vol, col="green")
abline(v=med_inf_vol, col="red")
abline(v=med_adu_vol, col="orange")
abline(v=smoothdiff_peak, col="black")
text(x=diff_peak, y=0, glue("Difference Peak: {round(smoothdiff_peak, 1)}"))

```

```{r 2Part_7c, fig.width = 12, fig.height = 6}

diff_df <- data.frame(Difference = difference, Smooth_difference = smooth.difference, Volume = volume.value)
diff_df <- melt(diff_df, id.vars="Volume", variable.name = "Type", value.name="Difference")
grid.arrange(
  ggplot(diff_df, aes(x = Volume, y = Difference, col = Type)) + 
    geom_line() + 
    geom_vline(xintercept = smoothdiff_peak, col='black') +
    geom_text(aes(x=smoothdiff_peak-20, label=glue("Smooth Difference Peak: {round(smoothdiff_peak, 1)}"), y=0), colour="black") +
    theme(legend.position = "bottom"),
  
  ggplot(data = harvest, aes(x = Volume, y = Proportion, col = Type)) +
    geom_line() +
    geom_vline(xintercept = max_inf_vol, col='green') + 
    geom_text(aes(x=max_inf_vol-10, label=glue("Protect All Infants: {round(max_inf_vol, 1)}"), y=0.5), colour="green", angle=90) +
    geom_vline(xintercept = med_inf_vol, col='red') + 
    geom_text(aes(x=med_inf_vol-10, label=glue("Median Infants: {round(med_inf_vol, 1)}"), y=0.5), colour="red", angle=90) +
    geom_vline(xintercept = med_adu_vol, col='orange') + 
    geom_text(aes(x=med_adu_vol-10, label=glue("Median Adults: {round(med_adu_vol, 1)}"), y=0.5), colour="orange", angle=90) +
    geom_vline(xintercept = smoothdiff_peak, col='black') +
    geom_text(aes(x=smoothdiff_peak-10, label=glue("Smooth Difference Peak: {round(smoothdiff_peak, 1)}"), y=0.5), colour="black", angle=90) +
    theme(legend.position = "bottom"),
  
    nrow=1
)

```


```{r 2Part_7d}

(1 - prop.adults)[which.max(smooth.difference)]  # [1] 0.7416332
# OR,
sum(mydata[mydata$TYPE == "ADULT", "VOLUME"] >
      volume.value[which.max(smooth.difference)]) / total.adults # [1] 0.7416332

(1 - prop.infants)[which.max(smooth.difference)]  # [1] 0.1764706
# OR,
sum(mydata[mydata$TYPE == "I", "VOLUME"] >
      volume.value[which.max(smooth.difference)]) / total.infants # [1] 0.1764706

```

-----

```{r 2Part_8a}

cutoff <- volume.value[volume.value > max(mydata[mydata$CLASS == "A1" &
  mydata$TYPE == "I", "VOLUME"])][1] # [1] 206.786

# Infant proportion
sum(mydata[mydata$TYPE == "I", "VOLUME"] > cutoff) / total.infants  # [1] 0.2871972 

# Adult proportion
sum(mydata[mydata$TYPE == "ADULT", "VOLUME"] > cutoff) / total.adults # [1] 0.8259705

```


```{r 2Part_8b, fig.width = 12}

ggplot(data = harvest, aes(x = Volume, y = Proportion, col = Type)) +
  geom_line() +
  geom_vline(xintercept = max_inf_vol, col='green') + 
  geom_text(aes(x=max_inf_vol-10, label=glue("Protect All Infants: {round(max_inf_vol, 1)}"), y=0.5), colour="green", angle=90) +
  geom_vline(xintercept = med_inf_vol, col='red') + 
  geom_text(aes(x=med_inf_vol-10, label=glue("Median Infants: {round(med_inf_vol, 1)}"), y=0.5), colour="red", angle=90) +
  geom_vline(xintercept = med_adu_vol, col='orange') + 
  geom_text(aes(x=med_adu_vol-10, label=glue("Median Adults: {round(med_adu_vol, 1)}"), y=0.5), colour="orange", angle=90) +
  geom_vline(xintercept = smoothdiff_peak, col='black') +
  geom_text(aes(x=smoothdiff_peak-10, label=glue("Smooth Difference Peak: {round(smoothdiff_peak, 1)}"), y=0.5), colour="black", angle=90) +
  geom_vline(xintercept = cutoff, col='darkblue') +
  geom_text(aes(x=cutoff-10, label=glue("Zero A1 infants: {round(cutoff, 1)}"), y=0.5), colour="darkblue", angle=90) +
  theme(legend.position = "bottom")
```

```{r 2Part_9, fig.width = 8.5}
x <- (1 - prop.infants)
y <- (1 - prop.adults)
max_inf_idx <- match(TRUE, volume.value>=max_inf_vol)
med_inf_idx <- match(TRUE, volume.value>=med_inf_vol)
med_adu_idx <- match(TRUE, volume.value>=med_adu_vol)
smd_peak_idx <- match(TRUE, volume.value>=smoothdiff_peak)
cutoff_idx <- match(TRUE, volume.value>=cutoff)
plot(x = x, y = y, type="l", xlab ="Infants Harvested", ylab="Adults Harvested")
abline(v=x[max_inf_idx], col="green")
text(x=x[max_inf_idx]-0.01, y=0.7, labels=glue("Protect All Infants: {round(max_inf_vol, 1)}"), col="green", srt=90)
abline(v=x[med_inf_idx], col="red")
text(x=x[med_inf_idx]-0.01, y=0.4, labels=glue("Median Infants: {round(med_inf_vol, 1)}"), col="red", srt=90)
abline(v=x[med_adu_idx], col="orange")
text(x=x[med_adu_idx]-0.01, y=0.7, labels=glue("Median Adults: {round(med_adu_vol, 1)}"), col="orange", srt=90)
abline(v=x[smd_peak_idx], col="black")
text(x=x[smd_peak_idx]-0.01, y=0.4, labels=glue("Difference Peak: {round(smoothdiff_peak, 1)}"), col="black", srt=90)
abline(v=x[cutoff_idx], col="darkblue")
text(x=x[cutoff_idx]-0.01, y=0.4, labels=glue("Zero A1 infants: {round(cutoff, 1)}"), col="darkblue", srt=90)
abline(v=x[cutoff_idx], col="darkblue")
abline(a=0, b=1, lty = 3)
```

```{r 2Part_9b}
auc(x = x, y = y)
```

```{r 2Part_10} 	
volume_cutoffs <- round(c(max_inf_vol, med_inf_vol, med_adu_vol, smoothdiff_peak, cutoff), 1)
cutoff_idxs <- c(max_inf_idx, med_inf_idx, med_adu_idx, smd_peak_idx, cutoff_idx)
round((1 - prop.adults)[cutoff_idxs[1]], 3)
x <- (1 - prop.infants)
y <- (1 - prop.adults)
sum(mydata$VOLUME > max_inf_vol) / length(mydata$VOLUME)

tpr <- c()
fpr <- c()
th <- c()
for (i in 1:length(volume_cutoffs)) {
  t <- round((1 - prop.adults)[cutoff_idxs[i]], 3)
  f <- round((1 - prop.infants)[cutoff_idxs[i]], 3)
  h <- round(sum(mydata$VOLUME > volume_cutoffs[i]) / length(mydata$VOLUME), 3)
  tpr <- c(tpr, t)
  fpr <- c(fpr, f)
  th <- c(th, h)
}

tab <- cbind(volume_cutoffs, tpr, fpr, th)
colnames(tab) <- c("Volume", "TPR", "FPR", "totalHarvest")
rownames(tab) <- c("Protect all infants", "Median infants", "Median adults", "Max difference", "Zero A1 infants")
tab <- as.table(tab)
tab
```

There are various thresholds depending on the cutoff volume and there are trade offs. Protect all infants is the most conservative threshold with only harvesting volumes that are equal or greater than 526.8. Because this method is protecting all infants, the FPR is 0, but TPR is limited to 0.246 as well. This will leave out many adult abalones that should have been harvested in order to protect all infants. The median infants, on the other hand, have a very aggressive cutoff volume at 133.8. This method has a high TPR in 0.933, but has a FPR of 0.498. This method harvests roughly half of infant abalones that shouldn't have been harvested in order to harvest more adults. The median adults, max difference, and zero A1 infants are in between the 2 extreme cutoff points. The business needs to make a decision based on the trade offs of harvesting the abalones now or keeping them for the future.
